Highly Available Transactions: Virtues and Limitations\\(Extended Version)

### 2013-05-01 writeup/conclusion.tex
Final submitted

HATs raise several challenges for future study. In this paper, we have
highlighted standardized models, but we would like a more concise
description of which models are and are not achievable in HAT
systems. Similarly, while we have studied the analytical and
experiment behavior of several HAT models, there is substantial work
in further understanding the performance and design of systems within
the large set of HAT models. We also believe it is time to revisit
techniques such as Sagas~\cite{sagas} and other long-running
transaction support in order to aid HAT programming. Weakened failure
assumptions as in escrow or in the form of bounded network asynchrony
could enable richer HAT semantics at the cost of general-purpose
availability. Alternatively, there is a range of possible hybrid
systems providing strong semantics during partition-free periods and
weakened semantics during partitions. Based on our understanding of
what is desired in the field and newfound knowledge of what is
possible to achieve, we believe HATs provide a large and useful design
space to explore.

>>>>

HATs raise several challenges for future study. In this paper, we have
highlighted standardized models, but we would like a more concise
description of which models are and are not achievable in HAT
systems. Similarly, while we have studied the analytical and
experiment behavior of several HAT models, there is substantial work
in further understanding the performance and design of systems within
the large set of HAT models. We also believe it is time to revisit
techniques such as Sagas~\cite{sagas} and other long-running
transaction support in order to aid HAT programming. Weakened failure
assumptions as in escrow or in the form of bounded network asynchrony
could enable richer HAT semantics at the cost of general-purpose
availability. Alternatively, there is a range of possible hybrid
systems providing strong semantics during partition-free periods and
weakened semantics during partitions. Based on our understanding of
what is desired in the field and newfound knowledge of what is
possible to achieve, we believe HATs provide a large and useful design
space for exploration.

### 2013-05-01 writeup/evaluation.tex
Final submitted

To further understand the performance implications of HAT guarantees
in a real-world environment, we implemented a HAT database
prototype. Across a range of deployments on public cloud
infrastructures, we found that, as Section~\ref{sec:latency}'s
measurements suggested, ``strongly consistent'' algorithms incur
substantial latency penalties: over WAN, 10 to 100 times higher than
their HAT counterparts. In light of recent algorithms for providing
causal consistency~\cite{bolton, cops, swift}, we focus our HAT
analysis on TA (algorithm from Section~\ref{sec:ta}, described in
greater in our Technical Report~\cite{hat-tr}), which we believe to be
the most novel of our algorithms. We find that HAT properties---while
more resource-intensive than basic eventual consistency---scale well
to many servers.

>>>>

To further understand the performance implications of HAT guarantees
in a real-world environment, we implemented a HAT database
prototype. Across a range of deployments on public cloud
infrastructures, we found that, as Section~\ref{sec:latency}'s
measurements suggested, ``strongly consistent'' algorithms incur
substantial latency penalties: over WAN, 10 to 100 times higher than
their HAT counterparts. Given recent algorithms for providing
causal consistency~\cite{bolton, cops, swift}, we focus our HAT
analysis on TA (algorithm from Section~\ref{sec:ta}, described in
greater detail
in Appendix B
%in our Technical Report~\cite{hat-tr}
), which we believe to be the
most interesting of our algorithms. We find that HAT
properties---while more resource-intensive than basic eventual
consistency---scale well to many servers.

### 2013-05-01 writeup/hat-analysis.tex
Final submitted

For brevity, we provide an informal presentation of each guarantee
here (accompanied by appropriate references) but give a full set of
formal definitions in our extended Technical Report~\cite{hat-tr}. In
our examples, we exclusively consider read and write operations,
denoting a write of value $v$ to data item $d$ as $w_d(v)$ and a read
from data item $d$ returning $v$ as $r_d(v)$. We assume that all data
items have the null value, $\bot$, at database initialization, and,
unless otherwise specified, all transactions in the examples commit.

@@

We sketch a highly available TA algorithm and provide greater detail
in our technical report~\cite{hat-tr}. We begin with our Read
Committed algorithm, but replicas wait to reveal new writes to readers
until all of the replicas for the final writes in the transaction have
received their respective writes (are \textit{pending
  stable}). Accordingly, replicas reveal writes asynchronously;
different replicas may detect pending stability at different times. To
prevent clients from reading incomplete portions of a transaction's
writes as they are being revealed (e.g., a client reads a data item
that from a replica that has detected pending stability, then attempts
to read another item from another replica that has not yet detected
pending stability), clients include additional metadata with each
write: a single timestamp for all writes in the transaction (e.g., as
in Read Uncommitted) and a list of items written to in the
transaction. When a client reads a write, the write's timestamp and
list of items form a lower bound on the versions that the client
should read for the other items. When a client reads any data item, it
attaches a timestamp to its request representing the current lower
bound for that item. Replicas use this timestamp to respond with
either a write matching the timestamp or a pending stable write with a
higher timestamp. Clients may discard any metadata upon transaction
commit, and servers keep two sets of writes for each data item: the
write with the highest timestamp that is pending stable and a set of
writes that are not yet pending stable.

>>>>

For brevity, we provide an informal presentation of each guarantee
here (accompanied by appropriate references) but give a full set of
formal definitions in
 Appendix A.
% our extended Technical Report~\cite{hat-tr}.
In our examples, we exclusively consider read and write operations,
denoting a write of value $v$ to data item $d$ as $w_d(v)$ and a read
from data item $d$ returning $v$ as $r_d(v)$. We assume that all data
items have the null value, $\bot$, at database initialization, and,
unless otherwise specified, all transactions in the examples commit.

@@

We sketch a highly available TA algorithm and provide greater detail
in Appendix B.
%our extended technical report~\cite{hat-tr}.
We begin with our Read Committed algorithm, but replicas wait to reveal new writes to readers
until all of the replicas for the final writes in the transaction have
received their respective writes (are \textit{pending
  stable}). Accordingly, replicas reveal writes asynchronously;
different replicas may detect pending stability at different times. To
prevent clients from reading incomplete portions of a transaction's
writes as they are being revealed (e.g., a client reads a data item
that from a replica that has detected pending stability, then attempts
to read another item from another replica that has not yet detected
pending stability), clients include additional metadata with each
write: a single timestamp for all writes in the transaction (e.g., as
in Read Uncommitted) and a list of items written to in the
transaction. When a client reads a write, the write's timestamp and
list of items form a lower bound on the versions that the client
should read for the other items. When a client reads any data item, it
attaches a timestamp to its request representing the current lower
bound for that item. Replicas use this timestamp to respond with
either a write matching the timestamp or a pending stable write with a
higher timestamp. Clients may discard any metadata upon transaction
commit, and servers keep two sets of writes for each data item: the
write with the highest timestamp that is pending stable and a set of
writes that are not yet pending stable.

### 2013-05-01 writeup/hat-vldb.tex
dat appendix
Final submitted

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
\usepackage{draftwatermarkbottom, draftwatermarktop}
\usepackage{scrextend, MnSymbol}

@@

\title{Highly Available Transactions: Virtues and Limitations}
{\author{Peter Bailis{\fontsize{12}{14}$^\dagger$}, Aaron Davidson{\fontsize{12}{14}$^\dagger$}, Alan Fekete{\fontsize{12}{14}$^\diamond$}, Ali Ghodsi{\fontsize{12}{14}$^{\dagger,\ddagger}$}, Joseph M. Hellerstein{\fontsize{12}{14}$^\dagger$}, Ion Stoica{\fontsize{12}{14}$^\dagger$}\\{\affaddr{\fontsize{12}{14}$^\dagger$}\hspace{.5mm}UC Berkeley\hspace{4mm}{\fontsize{12}{14}$^\diamond$}\hspace{.5mm}University of Sydney\hspace{4mm}{\fontsize{12}{14}$^\ddagger$}\hspace{.5mm}KTH/Royal Institute of Technology}}}
\maketitle

@@

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
%\pagebreak
%
\pagebreak
%\begin{appendix}
%\input{availability-appendix.tex}

\input{hat-appendix.tex}
%\end{appendix}
\end{document}

>>>>

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
%\usepackage{draftwatermarkbottom, draftwatermarktop}
\usepackage{scrextend, MnSymbol}

\usepackage[hidelinks]{hyperref}

@@

\title{Highly Available Transactions: Virtues and Limitations\\(Extended Version)}
{\author{Peter Bailis{\fontsize{12}{14}$^\dagger$}, Aaron Davidson{\fontsize{12}{14}$^\dagger$}, Alan Fekete{\fontsize{12}{14}$^\diamond$}, Ali Ghodsi{\fontsize{12}{14}$^{\dagger,\ddagger}$}, Joseph M. Hellerstein{\fontsize{12}{14}$^\dagger$}, Ion Stoica{\fontsize{12}{14}$^\dagger$}\\{\affaddr{\fontsize{12}{14}$^\dagger$}\hspace{.5mm}UC Berkeley\hspace{4mm}{\fontsize{12}{14}$^\diamond$}\hspace{.5mm}University of Sydney\hspace{4mm}{\fontsize{12}{14}$^\ddagger$}\hspace{.5mm}KTH/Royal Institute of Technology}}}
\maketitle

@@

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
\pagebreak

\pagebreak
\begin{appendix}

\input{hat-appendix.tex}
\end{appendix}
\end{document}
